Determining sequence-dependent DNA oligonucleotide hybridization and dehybridization mechanisms using coarse-grained molecular simulation, Markov state models, and infrared spectroscopy

### 2021-05-20 manuscript.tex
Andy final edits prior to submission

%\newcommand*{\addref}[1]{\textcolor{red}{\{ADD REF: #1\}}}
%\newcommand*{\noter}[1]{\textcolor{red}{[[#1]]}}		% notes 
%\newcommand*{\noter}[1]{}					% notes off
\newcommand*{\noteb}[1]{\textcolor{blue}{[[#1]]}}		% notes

@@

Our understanding of hybridization dynamics is built upon decades of experiments -- such as temperature-jump, salt-jump, pH-jump, and other perturbative methods -- that rapidly stimulate DNA and monitor relaxation to a new equilibrium.\citep{Morrison1993SensitiveSolution, Wetmur1968KineticsDNA, Craig1971RelaxationOligonucleotides, Porschke1973ThermodynamicsPairs, Williams1989LaserDGCATGC, Narayanan2012ExploringMixing, Chen2007InfluenceHybridization, Sanstead2018DirectDehybridization} More recently, single molecule diffusion and tethered multifluorophore assays have facilitated analyses under equilibrium conditions, but these results can be hampered by slow data collection rates and fluorescent tags effects on strand dynamics, particularly for shorter oligomers.\citep{Liu20173DSolution,  Schickinger2018TetheredHelices, Chen2008Base-by-baseSpectroscopy, Dupuis2013Single-moleculeHelices} A number of computational modeling approaches have also been employed to provide molecular-level resolution of hybridization. Simplified lattice models can recapitulate the essential aspects of the hybridization pathways but lack the realism of continuous space representations \citep{Araque2016LatticeCooperativity, Phys2019}. The long time scales associated with hybridization/dehybridization events place them outside the reach of unbiased all-atom molecular dynamics simulations,\citep{Hinckley2014Coarse-grainedEffects} but they can be observed by employing enhanced sampling techniques\citep{Piana2007AtomisticTransition, Zerze2021ThermodynamicsSimulations, Hinckley2013AnHybridization, Schmitt2013ExploringSurface, Sambriski2009, Sambriski2009SequencePathways, Hoefert2011MolecularOligonucleotides,Romano2013DNADependence, Hinckley2014Coarse-grainedEffects} or by using elevated temperature or denaturing solvent concentrations to induce one-way dissociation events.\citep{Wong2008TheSimulations, Perez2010Real-timeUnfolding} The effect of the applied bias upon the thermodynamics can be rigorously corrected for using standard reweighting techniques, but approaches to rigorously correct the kinetics, particularly under the conditions of high bias necessary for good sampling, are in their infancy.~\citep{Prinz2011OptimalDynamics, Chodera2011DynamicalTemperatures, Stelzl2017DynamicSimulations, Donati2017GirsanovModels, Donati2018GirsanovSimulations, Quer2018ANCOORDINATES} A number of coarse-grained DNA force fields have been developed that enable direct observation of these events over microsecond time scales via unbiased coarse-grained molecular dynamics simulations, \citep{Romano2013DNADependence, Hinckley2013AnHybridization, Maciejczyk2014DNAModel, Markegard2015, Dans2016MultiscaleDNA} which, up to a speedup factor associated with the smoothing of the underlying free energy landscape inherent to the coarse-graining procedure, can preserve a faithful model of the unbiased dynamics and associated pathways. These models have previously been used to study biological phenomena such as nucleosome dynamics\citep{Lequieu2016Tension-dependentUnwrapping, Lequieu2017InSliding} and transcription factor binding \cite{Terakawa2015P53Simulations, Tan2018DynamicData} as well as nanotechnology applications such as strand displacement \citep{Srinivas2013OnDisplacement, Haley2020DesignDisplacement} and DNA origami \citep{Snodin2019Coarse-grainedOrigami, Doye2020TheOrigami}.

In this work, we study a family of self-complementary pairs of 10-mer DNA oligomers using coarse-grained molecular simulation, machine learning of the slow dynamical modes, and data-driven inference of long-time kinetic models to establish new understanding of the influence of sequence upon hybridization/dehybridization kinetics and mechanisms. This family of oligomers -- 5$^\prime$-ATATATATAT-3$^\prime$ (AT-all), 5$^\prime$-GATATATATC-3$^\prime$ (GC-end), 5$^\prime$-ATATGCATAT-3$^\prime$ (GC-core), and 5$^\prime$-ATGATATCAT-3$^\prime$ (GC-mix) -- was designed to probe the influence of the placement of a G:C pair \noteb{``of two G:C base pairs?''}within an otherwise repetitive A:T sequence and has been the subject of our prior experimental investigations using temperature-jump infrared spectroscopy and simple lattice models. \citep{Sanstead2016} We validate the new computational models of hybridization/dehybridization dynamics developed in this work against new experimental data and reinterpret our prior experimental observations in light of the new computational understanding. Consistent with previous studies, \citep{Hinckley2014Coarse-grainedEffects,Romano2013DNADependence,Araque2016LatticeCooperativity} we find the degree of repetitiveness in the sequence -- and therefore the kinetic accessibility and thermodynamic stability of out-of-register shifted states -- leads to richer dynamics populated by a diversity of long-lived metastable states. Our data-driven modeling and analysis rigorously quantifies these behaviors and furnishes accurate predictive models of the hybridization/dehybridization rates, dynamical pathways, and metastable states. Specifically, we demonstrate that disrupting repetitive stretches of A:T bases by placement of interrupting G:C base pairs enables us to tune the landscape from rich six-state to simple two-state ``all-or-nothing'' behavior, and the specific location of the interrupting pair can be used to modulate the stability of long-lived frayed states. Taken together, our analyses establish new molecular-level understanding of the sequence-dependent kinetics and pathways through quantitative predictive models for the long-time system dynamics, resolution of the dynamical folding pathways and metastable states, and elementary design rules with which to sculpt and qualitatively alter the dynamical behaviors of the system. We anticipate that this new foundational understanding, and the extension of our approach to more extensive families of DNA sequences, can guide the rational design of DNA oligomers with tailored kinetic properties engineered for DNA nanotechnology applications such as DNA-PAINT. \citep{Shah2019, Strauss2020UpDNA-PAINT}

% example text from Andy's conclusions:

% Correlation of the data-driven kinetic order parameters extracted by SRVs provides new understanding of the sequence-dependent kinetic mechanisms by which oligomers hybridize and reveal these to be richer than classical base-pair shifting and “all-or-nothing” behaviors.

% original last paragraph
%By evaluating equilibrium trajectories we can study the relevance of metastable states during both the hybridization and dissociation process. Because SRVs generate an optimized low dimensional basis, we show that higher resolution MSMs (measured by a reduction in the required lag time) are accessible. Additionally, we compare slow dynamical modes and metastable states between sequence-specific SRV-MSMs. Within these metastable dynamical states, we leverage diffusion maps to analyze the diversity of structures whose inter-conversion rate are too fast to produce unique slow modes. Finally, we run higher temperatures simulations to investigate the temperature-dependent nature of some experimentally relevant dynamical responses. Taken together, our analysis reflects similar results to previous computational and experimental DNA work, while elucidating new insights into sequence-dependent dynamics, metastable structures, and relative time scales.

%These studies have fueled increasingly complex structural DNA technologies, but developments in dynamic DNA nanotechnologies have outpaced our understanding of the dynamics themselves \citep{Deluca2020DynamicDevices}.

%Still, rich dynamics have been observed short nucleotides \citep{Sanstead2018DirectDehybridization} and the sequence-dependent effects on these dynamics are not well understood.

% include Tokmakoff conclusions in the intro -- sequence-dependent deviations from two-state behavior. The slow response was attributed to complete strand dissociation, while the fast response described base fraying. We perform longer equilibrium sampling near Tm to get a sense of sequence equilibrium behavior, as well shorter simulations at a range of higher temperatures to replicate T-jump data. Taken together, our analysis reflects similar results to previous computational and experimental DNA work, while elucidating new insights into sequence-dependent dynamics, metastable structures, and relative time scales.

@@

%\subsection{Featurization}

%A featurization of trajectories was required to capture fundamental dynamics while accounting for translational and rotational symmetries. Pairwise distances between base centers of mass -- similar to heavy atom distances in protein models \citep{Sidky}-- satisfied these requirements and were calculated at each frame using the MDtraj Python library.\citep{McGibbon2015MDTraj:Trajectories} Although pairwise distances remove translational and rotational symmetries, additional symmetries arise from the self-complementary nature of these sequences -- the sense and anti-sense strands in each pair are identical to each other. To account for this, we averaged permutable distances (45 pairs in total) together into a ``symmetrized'' feature set following a similar procedure used in tICAgg coordinate construction.\citep{Sengupta2019AutomatedSelf-assembly} The VAMP-2 scoring method was employed to evaluate the kinetic variance of the feature set and optimize hyperparameters.\cite{ Mardt2018VAMPnetsKinetics} Additional details on the VAMP-2 score can be found in the supplemental information. By comparing scores across several candidate feature sets (Figure \ref{fig:allseq_features_vamp2}), we found that minimal kinetic information was lost in the  symmetrization procedure. Furthermore, the smaller feature set enabled faster training times and better statistics over permutable distances, without appearing to suffer a loss in generality or model resolution.

%\begin{figure}[ht!]
%	\begin{center}
%        \includegraphics[width=0.55\textwidth]{Fig1.pdf}
%        \caption{A comparison of VAMP-2 scores via 5-fold cross-validation for the intermolecular, intra+intermolecular,  and symmetrized intermolecular distances, in addition to the reciprocal value for each of these coordinates. Reciprocal coordinates performed better across the board, but there was no substantial difference between the three featurization methods. This indicated that minimal kinetic information was lost in the symmetrization procedure, therefore we chose reciprocal symmetrized coordinates to train the model.}
%        \label{fig:allseq_features_vamp2}
%	\end{center}
%\end{figure}

%\subsection{\label{sec:methods}SRVs}

%The next step in the MSM pipeline is to create a low-dimensional kinetic projection where state clustering is performed. While most pipelines employ tICA to generate a linear transformation of input features, we use SRVs as a more powerful modular replacement for tICA. SRVs were first developed by Chen et al. as a means to directly learn slow eigenfunctions of the transfer operator.\citep{Chen} The method is a descendant of VAMPnets\citep{Mardt2018VAMPnetsKinetics}, deep canonical correlation analysis \citep{Andrew2013DeepAnalysis}, and extended dynamic mode decomposition with dictionary learning (EDMD-DL).\citep{Li2017ExtendedOperator} The framework uses a twin-lobed artificial neural network to learn an optimal basis set for the variational approach to conformational dynamics (VAC) from which the leading eigenfunctions of the transfer operator are then estimated.\citep{Noe2013ASystems} The resulting orthogonal modes are associated with the slowest dynamical processes in a system, and can be used to interpret kinetic information directly (such as physical correlations and time scales) and to construct MSMs.\citep{Sidky, Pande2010EverythingAsk} SRVs provide robust nonlinear approximation and computation time that scales linearly with the amount of input data and are more powerful and efficient that tICA and ktICA approaches. This is a key attribute to our system as 10 million frames with 55 features per frame are used for each sequence. The SRV framework has been used to construct high kinetic resolution MSMs applied to molecular simulations of WW-domain and Trp-cage mini-protein.\citep{Chen, Sidky}

%\subsection{\label{sec:methods}SRV-MSMs}

% need some more detail here -- talk to Andy about this
%MSMs are a powerful tool for interpreting large amounts of simulation data in a statistically robust and experimentally comparable way.\citep{Prinz2011OptimalDynamics, Husic2018MarkovScience} Kinetically similar conformations are first discretized into microstates, and the conditional probability of transitions between microstates is calculated within some lag time. The reliance on conditional probabilities allows for many independent simulations (longer than the lag time) to be collectively interpreted. To take full advantage of the MSM frameworks, however, the input basis should be as kinetically meaningful as possible.\citep{Pande2010EverythingAsk} Because SRV eigenfunctions translate simulation features into their slowest kinetic representations, they function as an optimal MSM basis. To build our SRV-MSM framework, we employed the PyEmma MSM pipeline and generated independent models for each sequence.\citep{Scherer2015PyEMMAModels} In a similar approach to Sidky et al. we performed k-means microstate clustering, Bayesian MSM construction, and PCCA+ hierarchical macrostate assignments.\citep{Sidky} The number of microstates were determined by VAMP-2 score, and the SRV-MSM lag time was selected based on implied time scales convergence.  The number of PCCA+ macrostates was determined based on the characteristic of each system and is discussed more in depth in the results and supplemental information.

%\subsection{\label{sec:methods}Diffusion Maps}
%Diffusion maps generate a low-dimensional non-linear embedding of data into high variance structural modes.\citep{Coifman2006DiffusionMaps, Ferguson2010SystematicMaps} They are well-suited to find subtle differences in molecular conformations that are temporally disconnected or that interchange faster than the MSM lag time. We used reciprocal inter-molecular base distances as inputs to diffusion maps, and tuned $\epsilon$ and $\alpha$ parameters in accordance with previous studies.\citep{Shmilovich, Wang2017MesoscaleLandscapes} We found that the choice of these parameters mainly served to condense or stretch the low dimensional space and had a minimal qualitative impact on salient results.

%%%%%%%% Methods comments %%%%%%%%%%%%%%%%%

%We found close agreement with explicit ion simulations performed with 240 mM \ce{NaCl} and 18 mM \ce{MgCl2} (supplemental) but 7x slower simulation speed made adequate sampling more difficult to achieve under these conditions.

%We initialized explicit ions such that 240 mM \ce{NaCl} and 18 mM \ce{MgCl2} were added to the box in addition to 18 Na counter ions to balance the charge from the 9 phosphate groups in each oligonucleotide backbone \citep{Hinckley2015}. We used a periodic box size of 7.774 nm and an effective oligo concentration of 7 mM. We used an Ewald summation to calculate long range Coulombic interaction between DNA and ions using a real space cutoff of 2 nm.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

% rerun these at the actual final temps and just see what happens? Doesn't really matter where Tjump occurs and could make things easier as a whole...
% talk about rerunning AT-all and GC-mix=

@@

To computationally mimic the T-jump process in our 3SPN.2 simulations, we conducted 1 $\mu$s simulations of an initially hybridized DNA duplex over a range of temperatures and monitored its structural relaxation. We performed 120 independent simulations for each DNA sequence at each temperature, and from these extracted computational estimates of $k_d^\mathrm{slow}$ and fast $k_d^\mathrm{fast}$ (\blauw{Fig.~S4}). First, we tracked the slow response corresponding to duplex dissociation in our simulations by compiling the distribution of times at which both of the central base pairs first separate to a distance of 1.3 nm starting from an initial fully hybridized duplex. This cutoff was selected as the distance beyond with the strands are effectively non-interacting and defines the dissociated state. We extracted our computational estimate of $k_d^\mathrm{slow}$ by fitting a decaying exponential to the fraction of hybridized sequences as a function of time $f_\mathrm{hybridized}(t) = \exp(-k_d^\mathrm{slow}t)$. Second, we tracked the fast response corresponding to terminal base pair fraying by compiling the distribution of times at which either of the terminal base pairs first separated to a distance of 1.3 nm, corresponding to a complete breakage of the WC interaction. We extracted our computational estimate of $k_d^\mathrm{fast}$ through a decaying exponential fit to the fraction of unfrayed sequences as a function of time $f_\mathrm{unfrayed}(t) = \exp(-k_d^\mathrm{fast}t)$.

We present in Fig.~\ref{fig:relaxation-comparison} a comparison of $k_d^\mathrm{slow}$ and $k_d^\mathrm{fast}$ estimated by computation and experiment. Although the 3SPN.2 model was not directly fitted against kinetic data,~\cite{Hinckley2013AnHybridization} its predictions of sequence dependent T-jump relaxation rates are, within a systematic scaling factor in time and systematic shift in temperature, in good agreement with observed experimental trends. It is well-known that the smoothing of the underlying free energy landscape induced by coarse-graining artificially accelerates the kinetics of coarse-grained molecular simulations and that different degrees of freedom may be accelerated by different factors.~\citep{Marrink2007TheSimulations, Fritz2011MultiscaleDynamics, Marrink2013PerspectiveModel} We find that the simulated slow responses corresponding to center-of-mass translation of the strands during dissociation of the duplex is $\sim$10$\times$ accelerated relative to experiment, whereas the fast responses corresponding to fraying of the terminal bases is $\sim$120$\times$ accelerated. We apply these sequence-independent scaling factors to our reported computational values \noteb{I know we clarify this at the end of the paragraph, but maybe add ``our reported computational values in Fig.~\ref{fig:relaxation-comparison}'' to be clear?}. Although the 3SPN.2 model reproduces melting temperatures relatively well, we observed a systematic 4 K under-prediction relative to experiment and so we apply a universal (+4) K corrective calibration to our computational results. We note that these empirical calibration factors to the 3SPN.2 predictions are applied only for the purposes of making an experimental comparison, but acknowledge that there are uncertainties introduced by assuming that the equilibrium dynamics at fixed temperature can be compared directly to relaxation kinetics following a 15$^\circ$C T-jump. The computational time scales and melting temperatures reported in the remainder of the paper are not corrected since we are only interested in the relative trends in the behaviors of the four sequences.

@@

In Fig.~\ref{fig:relaxation-comparison}a we observe an exponential increase of $k_d^\mathrm{slow}$ with temperature, as expected from the large enthalpic barrier to duplex dissociation.\cite{Craig1971RelaxationOligonucleotides,Porschke1971CooperativeTransition,Williams1989LaserDGCATGC} Under the time and temperature calibration corrections, we see generally very good agreement between the computational and experimental curves. The computational rates do tend to systematically under-predict the experimental values at low temperatures, an effect that we ascribe to the fact that the experimental T-jump responses likely contain a mixture of dissociation and hybridization events -- the latter of which persist to lower temperatures than the former -- whereas the computational analysis considers dissociation alone \noteb{This effect is now minimized by the fact that we are using an experimental dissociation rates instead of the pure slow response (previous data showed worse agreement at low temp). Still, there could be some hybridization contribution so we could leave this sentence in}. Of the four sequences, GC-core shows the largest discrepancy between computation and experiment, although the general exponential trend is preserved. This may be a result of its high propensity to fray (cf.\ Section \ref{sec:GC-core}). In Fig.~\ref{fig:relaxation-comparison}b we expose a largely linear dependence of $k_d^\mathrm{fast}$ upon temperature for AT-all, GC-end, and GC-mix compared to an exponential dependence for GC-end. These trends can be understood in light of the comparatively larger enthalpic barrier for dissociation of the terminal G:C base pair in GC-end compared to that for the A:T terminal pair in the other three sequences. Again we see good agreement between the scaled computational predictions and the experimental T-jump measurements. The favorable comparison of computation and experiment provides support for the capacity of the 3SPN.2 model to reliably reproduce sequence-dependent trends in the slow and fast kinetics of the four DNA oligomers.

%Coarse-grained time scales are inherently difficult to match against experiments due to smoothing of the free energy landscape and variable acceleration across different degrees of freedom. As such, we evaluated multiple temperatures for each sequence and inferred acceleration factors from temperature-dependent trends.  Experimental temperature-dependent relaxation time scales were measured for each sequence using T-jump IR spectroscopy as described previously.\citep{Sanstead2018DirectDehybridization} We recorded ``fast'' and ``slow'' amplitude-weighted responses -- previously attributed to terminal base pair fraying and duplex dissociation, respectively -- for each temperature and sequence. To compare results with 3SPN.2, we ran 120 shorter (1 us) simulations initialized in the hybridized state at final T-jump temperatures for each sequence. We calculated relaxation times for a slow dissociation response and fast fraying response at each temperature and compared these with experimental temperature-dependent relaxation fits.

% need to discuss tmelt shift
%We measured the slow dissociation response by fitting the distribution of times at which the core base pairs separate beyond a 1.3 nm cutoff. We found the inverse of these relaxation times -- the effective dissociation rate -- to increase exponentially with temperature, which is expected given the large enthalpic barrier of dissociation.\citep{Craig1971RelaxationOligonucleotides,Porschke1971CooperativeTransition,Williams1989LaserDGCATGC} We noted an average temperature shift of about 4$^{\circ}$C between experiment and simulation fits. While 3SPN.2 captures melting temperature relatively well, some deviation is expected given varied ion conditions and other coarse-grained effects. We also noted an acceleration of about one order of magnitude compared to experiment, although this factor is sensitive to the exact definition of melting temperature. After accounting for the melting temperature shift and acceleration factor, we saw strong agreement between the experimental data and simulated relaxation fits (Figure \ref{fig:relaxation-comparison})). At lower temperatures, T-jump slow responses likely contain a mixture of dissociation and hybridization dynamics, therefore rates do not drop off to the same extent as our analysis, which considers dissociation alone. GC-core showed the largest deviation from experiment, indicating that increased temperature triggers dissociation at a disproportionately higher rate in simulation compared to experiment.

% GC-core fast time scales could be slower as they have contributions from internal fraying which would have a longer relaxation than the termini alone.
% in order to make comparisons across these different free energy landscapes, we adjusted simulations by a consistent acceleration factor. We maintained a consistent accegqrelation factor across sequences, but found that optical scaling different by about one order of magnitude between fast and slow responses. It is expected given that different degrees of freedom often experience differing amounts of accelerations in coarse-grained models.

% need to update this fig with labels and temp shift

% A documented short-coming for the 3SPN.2 model is a high proclivity for termini fraying relative to what has been observed in vitro \citep{2013}. This supports why we observe significantly more acceleration for the fast response, but relative sequence and temperature effects still compare well with T-jump results.

%The fast response -- which Sanstead et. al attributed to terminal base pair fraying signatures -- was more challenging to compare against simulation observables. Numerous experimental and computational studies have shown that DNA and RNA fraying is a complex dynamical process with time scales that span 5 ps to several microseconds.\citep{ Nonin1995TerminalFraying, Nikolova2012ProbingSimulations, Andreatta2006UltrafastHelix, Galindo-Murillo2015ConvergenceDGCACGAACGAACGAACGC} All-atom simulations suggest that frayed ends can assume misaligned WC bonds, base-sugar hydrogen bonds, and terminal stacked conformations.\citep{PinamontiTheModels, Zgarbova2014BaseRNA} Given that there is only one interaction site parameterized on each 3SPN.2 base, we would not expect to resolve this diverse collection of states and dynamics. Instead, we measured the fast fraying response by counting frames until duplex terminal ends splits beyond a cutoff. This approach assumes that fraying on the permutable top and bottom of the duplex are independent from each other, and that a base pair distance is a reliable approximation for the ensemble spectroscopy signal. This is a reasonable assumption given that the amplitude-weighted time scales should consist largely of terminal fraying events. Again we fit relaxation curves to the ensemble of fraying time scales in order to extract rate approximations for each sequence at a series of temperatures.

%For A:T terminal sequences, the simulated fast response appears linear with temperature, indicating a barrierless and diffusion-driven process. In contrast, the GC-end responses are distinctly slower and increase exponentially with temperature, likely due to a greater enthalpic barrier associated with G:C fraying. We observe similar trends in the experimental data, although comparisons at the highest temperatures were limited due to mixing between the fast and slow responses (Figure \ref{fig:relaxation-comparison}). We found the optimal acceleration factor to be dependent on the choice of the cutoff parameter, but we can approximate that 3SPN.2 fraying dynamics are accelerated by about 100-200x relative to experiment. It is not surprising that we see different rates of acceleration for the dissociation and fraying processes given that coarse-grained effects can vary across different degrees of freedom. In particular, the simplified treatment of the fraying process may smoothen the free energy landscape and speed up dynamics relative to a more global processes like dissociation. Although one should rely on higher resolution models to study in depth mechanisms of fraying, our results indicate that terminal fraying is a reasonable assignment for the spectroscopic fast response in Sanstead et al. and that 3SPN.2 can recapitulate some sequence-dependent fraying effects.
% need a better transition here back into SRV-MSMs?

@@

%We followed the SRV-MSM analysis pipeline described in the Methods section to generate sequence-dependent kinetic models for each oligomer. We identified seven kinetically relevant states that were captured within the resolution of the model. Figure \ref{fig:allseq_table} shows cartoon and molecular renderings for each of these states as well as their sequence-specific stationary probabilities.  Figure \ref{fig:allseq_transition} shows how these states fit into sequence-specific low-dimensional representations and kinetic models. Arrows between states show the probability of a transition to another state within the MSM lag time.

%Shifted states are only observed for AT-all and GC-end; state abbreviations indicate the direction of the shifted overhang (5$^\prime$ vs. 3$^\prime$) and the number of shifted motifs relative the native state (1 vs. 2). GC-mix displays two-state behavior within the resolution of the model, however we will show that the hybridization and dissociation are still characterized by an ensemble of shorter-lived states. Stationary probabilities in each of these states are shown in figure \ref{fig:allseq_table}. We also show that PCCA+ clustering assigns states to relative free energy minima in the tICA space.

% rewrite this into one paragraph

% change state names and scientific notation
% include D here? can see what it would look like...

@@

%Given that terminal mismatches tend to be slightly stabilizing \citep{Santalucia2004TM}, we might expect a lower discrepancy than the 10.1-14.7 kJ/mol offset that we observe between NN and MSM results. This offset may arise in part from the excluded volume interactions used to account for mismatched bases in 3SPN.2, however we attribute it mainly to inert tail effects.

@@

%\bood{As illustrated in Fig.~\ref{fig:NN_table_no_de}, we see that our macrostate free energies recapitulate NN model predictions reasonably well. It is important to note that we do not include dangling end contributions when calculating NN free energies. Although 3SPN.2 can capture some dangling end stabilization effects through base stacking and cross-stacking interactions, the CG model was not parameterized to fully capture the enthalpic contributions of free bases interaction with terminal base pairs. For each macrostate, dangling ends would be expected to have stabilizing effects (Can show NN with dangling end contributions in SI and therefore increase state populations compared to what we observe in simulation.)

%\bood{A well known deficiency of the NN model is the absence of any treatment of inert tails -- free bases that extend beyond the dangling end are known to destabilize the duplex. We find that inert tail effects may be particularly impact in predicting out-of-register state stability.}

>>>>

%\newcommand*{\addref}[1]{\textcolor{red}{\{ADD REF: #1\}}}
%\newcommand*{\noter}[1]{\textcolor{red}{[[#1]]}}		% notes 
%\newcommand*{\noter}[1]{}					% notes off
%\newcommand*{\noteb}[1]{\textcolor{blue}{[[#1]]}}		% notes

@@

Our understanding of hybridization dynamics is built upon decades of experiments -- such as temperature-jump, salt-jump, pH-jump, and other perturbative methods -- that rapidly stimulate DNA and monitor relaxation to a new equilibrium.\citep{Morrison1993SensitiveSolution, Wetmur1968KineticsDNA, Craig1971RelaxationOligonucleotides, Porschke1973ThermodynamicsPairs, Williams1989LaserDGCATGC, Narayanan2012ExploringMixing, Chen2007InfluenceHybridization, Sanstead2018DirectDehybridization} More recently, single molecule diffusion and tethered multifluorophore assays have facilitated analyses under equilibrium conditions, but these results can be hampered by slow data collection rates and fluorescent tags effects on strand dynamics, particularly for shorter oligomers.\citep{Liu20173DSolution,  Schickinger2018TetheredHelices, Chen2008Base-by-baseSpectroscopy, Dupuis2013Single-moleculeHelices} A number of computational modeling approaches have also been employed to provide molecular-level resolution of hybridization. Simplified lattice models can recapitulate the essential aspects of the hybridization pathways but lack the realism of continuous space representations \citep{Araque2016LatticeCooperativity, Phys2019}. The long time scales associated with hybridization/dehybridization events place them outside the reach of unbiased all-atom molecular dynamics simulations,\citep{Hinckley2014Coarse-grainedEffects} but they can be observed by employing enhanced sampling techniques\citep{Piana2007AtomisticTransition, Zerze2021ThermodynamicsSimulations, Hinckley2013AnHybridization, Schmitt2013ExploringSurface, Sambriski2009, Sambriski2009SequencePathways, Hoefert2011MolecularOligonucleotides,Romano2013DNADependence, Hinckley2014Coarse-grainedEffects} or by using elevated temperature or denaturing solvent concentrations to induce one-way dissociation events.\citep{Wong2008TheSimulations, Perez2010Real-timeUnfolding} The effect of the applied bias upon the thermodynamics can be rigorously corrected for using standard reweighting techniques, but approaches to rigorously correct the kinetics, particularly under the conditions of high bias necessary for good sampling, are in their infancy.~\citep{Prinz2011OptimalDynamics, Chodera2011DynamicalTemperatures, Stelzl2017DynamicSimulations, Donati2017GirsanovModels, Donati2018GirsanovSimulations, Quer2018ANCOORDINATES} A number of coarse-grained DNA force fields have been developed that enable direct observation of these events over microsecond time scales via unbiased coarse-grained molecular dynamics simulations, \citep{Romano2013DNADependence, Hinckley2013AnHybridization, Maciejczyk2014DNAModel, Markegard2015, Dans2016MultiscaleDNA} which, up to an acceleration factor associated with the smoothing of the underlying free energy landscape inherent to the coarse-graining procedure, can preserve a faithful model of the unbiased dynamics and associated pathways. These models have previously been used to study biological phenomena such as nucleosome dynamics\citep{Lequieu2016Tension-dependentUnwrapping, Lequieu2017InSliding} and transcription factor binding \cite{Terakawa2015P53Simulations, Tan2018DynamicData} as well as nanotechnology applications such as strand displacement \citep{Srinivas2013OnDisplacement, Haley2020DesignDisplacement} and DNA origami \citep{Snodin2019Coarse-grainedOrigami, Doye2020TheOrigami}.

In this work, we study a family of self-complementary pairs of 10-mer DNA oligomers using coarse-grained molecular simulation, machine learning of the slow dynamical modes, and data-driven inference of long-time kinetic models to establish new understanding of the influence of sequence upon hybridization/dehybridization kinetics and mechanisms. This family -- 5$^\prime$-ATATATATAT-3$^\prime$ (AT-all), 5$^\prime$-GATATATATC-3$^\prime$ (GC-end), 5$^\prime$-ATATGCATAT-3$^\prime$ (GC-core), and 5$^\prime$-ATGATATCAT-3$^\prime$ (GC-mix) -- was designed to probe the influence of the placement of two G:C base pairs within an otherwise repetitive A:T sequence and has been the subject of our prior experimental investigations using temperature-jump infrared spectroscopy and simple lattice models. \citep{Sanstead2016} We validate the new computational models of hybridization/dehybridization dynamics developed in this work against new experimental data and reinterpret our prior experimental observations in light of the new computational understanding. Consistent with previous studies, \citep{Hinckley2014Coarse-grainedEffects,Romano2013DNADependence,Araque2016LatticeCooperativity} we find the degree of repetitiveness in the sequence -- and therefore the kinetic accessibility and thermodynamic stability of out-of-register shifted states -- leads to richer dynamics populated by a diversity of long-lived metastable states. Our data-driven modeling and analysis rigorously quantifies these behaviors and furnishes accurate predictive models of the hybridization/dehybridization rates, dynamical pathways, and metastable states. Specifically, we demonstrate that disrupting repetitive stretches of A:T bases by placement of interrupting G:C base pairs enables us to tune the landscape from rich six-state to simple two-state ``all-or-nothing'' behavior, and the specific location of the interrupting pair can be used to modulate the stability of long-lived frayed states. Taken together, our analyses establish new molecular-level understanding of the sequence-dependent kinetics and pathways through quantitative predictive models for the long-time system dynamics, resolution of the dynamical folding pathways and metastable states, and elementary design rules with which to control the dynamical behaviors of the system. We anticipate that this new foundational understanding, and the extension of our approach to more extensive families of DNA sequences, can guide the rational design of DNA oligomers with tailored kinetic properties engineered for DNA nanotechnology applications such as DNA-PAINT. \citep{Shah2019, Strauss2020UpDNA-PAINT}

@@

@@

@@

@@

To computationally mimic the T-jump process in our 3SPN.2 simulations, we conducted 1 $\mu$s simulations of an initially hybridized DNA duplex over a range of temperatures and monitored its structural relaxation. We performed 120 independent simulations for each DNA sequence at each temperature, and from these extracted computational estimates of $k_d^\mathrm{slow}$ and $k_d^\mathrm{fast}$ (\blauw{Fig.~S4}). First, we tracked the slow response corresponding to duplex dissociation in our simulations by compiling the distribution of times at which both of the central base pairs first separate to a distance of 1.3 nm starting from an initial fully hybridized duplex. This cutoff was selected as the distance beyond with the strands are effectively non-interacting and defines the dissociated state. We extracted our computational estimate of $k_d^\mathrm{slow}$ by fitting a decaying exponential to the fraction of hybridized sequences as a function of time $f_\mathrm{hybridized}(t) = \exp(-k_d^\mathrm{slow}t)$. Second, we tracked the fast response corresponding to terminal base pair fraying by compiling the distribution of times at which either of the terminal base pairs first separated to a distance of 1.3 nm, corresponding to a complete breakage of the WC interaction. We extracted our computational estimate of $k_d^\mathrm{fast}$ through a decaying exponential fit to the fraction of unfrayed sequences as a function of time $f_\mathrm{unfrayed}(t) = \exp(-k_d^\mathrm{fast}t)$.

We present in Fig.~\ref{fig:relaxation-comparison} a comparison of $k_d^\mathrm{slow}$ and $k_d^\mathrm{fast}$ estimated by computation and experiment. Although the 3SPN.2 model was not directly fitted against kinetic data,~\cite{Hinckley2013AnHybridization} its predictions of sequence dependent T-jump relaxation rates are, within a systematic scaling factor in time and systematic shift in temperature, in good agreement with observed experimental trends. It is well-known that the smoothing of the underlying free energy landscape induced by coarse-graining artificially accelerates the kinetics of coarse-grained molecular simulations and that different degrees of freedom may be accelerated by different factors.~\citep{Marrink2007TheSimulations, Fritz2011MultiscaleDynamics, Marrink2013PerspectiveModel} We find that the simulated slow responses corresponding to center-of-mass translation of the strands during dissociation of the duplex is $\sim$10$\times$ accelerated relative to experiment, whereas the fast responses corresponding to fraying of the terminal bases is $\sim$120$\times$ accelerated. We apply these sequence-independent scaling factors to our reported computational values in Fig.~\ref{fig:relaxation-comparison}. Although the 3SPN.2 model reproduces melting temperatures relatively well, we observed a systematic 4 K under-prediction relative to experiment and so we apply a universal (+4) K corrective calibration to our computational results. We note that these empirical calibration factors to the 3SPN.2 predictions are applied only for the purposes of making an experimental comparison, but acknowledge that there are uncertainties introduced by assuming that the equilibrium dynamics at fixed temperature can be compared directly to relaxation kinetics following a 15$^\circ$C T-jump. The computational time scales and melting temperatures reported in the remainder of the paper are not corrected since we are only interested in the relative trends in the behaviors of the four sequences.

@@

In Fig.~\ref{fig:relaxation-comparison}a we observe an exponential increase of $k_d^\mathrm{slow}$ with temperature, as expected from the large enthalpic barrier to duplex dissociation.\cite{Craig1971RelaxationOligonucleotides,Porschke1971CooperativeTransition,Williams1989LaserDGCATGC} Under the time and temperature calibration corrections, we see generally very good agreement between the computational and experimental curves. Of the four sequences, GC-core shows the largest discrepancy between computation and experiment, although the general exponential trend is preserved. This may be a result of its high propensity to fray (cf.\ Section \ref{sec:GC-core}). In Fig.~\ref{fig:relaxation-comparison}b we expose a largely linear dependence of $k_d^\mathrm{fast}$ upon temperature for AT-all, GC-end, and GC-mix compared to an exponential dependence for GC-end. These trends can be understood in light of the comparatively larger enthalpic barrier for dissociation of the terminal G:C base pair in GC-end compared to that for the A:T terminal pair in the other three sequences. Again we see good agreement between the scaled computational predictions and the experimental T-jump measurements. The favorable comparison of computation and experiment provides support for the capacity of the 3SPN.2 model to reliably reproduce sequence-dependent trends in the slow and fast kinetics of the four DNA oligomers.

@@

@@

@@
